---
title: "Stats Down South (Park)"
---

## Summary

My objective was to predict the average rating of a South Park episode on IMDb. Ultimately, I found that the k-nearest neighbors model I came up with doesn't have any practical uses, since both the response and the predictors are equally accessible on the same website.

Furthermore, since the data on IMDb comes from a voluntary sample, the response variable (the average rating of an episode) tended to be very high. This can be seen in the histogram below showing data from the training set, where the x-axis shows the average rating out of 10 stars and the heights of the bars show the number of episodes with that average rating. We can see that the histogram is centered at about 8 stars, with only a few episodes getting an average rating below 7 stars.

```{r}
#| label: AverageRating histogram for summary
#| message: FALSE
#| echo: FALSE

ggplot2::ggplot(
  data = SouthParkRshiny::SouthPark_IMDB_Data[which(SouthParkRshiny::SouthPark_IMDB_Data$Year <= 2013),],
  mapping = ggplot2::aes(
    x = AverageRating
  )
) +
  ggplot2::geom_histogram() +
  ggplot2::labs(
    x = "Average Rating out of 10",
    y = "Number of Episodes",
    title = "South Park IMDb"
  )
```

Because of this, the model tended to overestimate episodes with a lower average rating.

## Motivation and Context

```{r}
#| label: do this first
#| echo: false
#| message: false

here::i_am("math-437-project.qmd")
```

I am investigating whether or not the average rating of a South Park episode on IMDb can be predicted using data about the episode such as the number of votes it got on IMDb. Specifically, what I am trying to predict is the average of the ratings an episode received among all of the users that gave it a rating from 1 to 10 stars. The reason I chose to investigate this is because the average rating of an episode seemed like the most appropriate metric for measuring how well an episode did among audiences, which can be useful in determining how future episodes should be written.

## Packages Used In This Analysis

```{r}
#| label: load packages
#| message: false
#| warning: false

library(here)
library(dplyr)
library(ggplot2)
library(parsnip)
library(recipes)
library(workflows)
library(rsample)
library(tune)
library(yardstick)
```

| Package | Use |
|-------------------------------|----------------------------------------|
| [here](https://github.com/jennybc/here_here) | to easily load and save data |
| [dplyr](https://dplyr.tidyverse.org/) | to massage and summarize data |
| [ggplot2](https://ggplot2.tidyverse.org/) | to create nice-looking and informative graphs |
| [parsnip](https://parsnip.tidymodels.org/) | to create the model |
| [recipes](https://recipes.tidymodels.org/) | to prepare data for modeling |
| [workflows](https://workflows.tidymodels.org/) | to put together and finalize the model |
| [rsample](https://rsample.tidymodels.org/) | to set up cross-validation |
| [tune](https://tune.tidymodels.org/) | to perform cross-validation |
| [yardstick](https://yardstick.tidymodels.org/) | to set metrics for cross-validation |

## Data Description

I will be using data from the `SouthPark_IMDB_Data` data set. Information about the data can be found [here](https://github.com/pdrhlik/southparkr). This data was collected by Amalan Mahendran, or Amalan-ConStat on GitHub. The information in the data set was taken from South Park's IMDb website and seems to have been collected to be used in other users' statistical analysis. One observation represents one episode of South Park. There are 326 observations and 8 variables in this data set.

```{r}
#| label: import data
#| message: false
#| warning: false

library(SouthParkRshiny)
```

## Data Wrangling

I split the data by putting episodes from 2013 or earlier into the training set, and episodes from after 2013 into the test set. This left me with a reasonable amount of data in both sets (248 observations in the training set and 78 observations in the test set). I did the split based on the year the episodes were aired because this puts the earliest episodes in the training set and the latest episodes in the test set, which prevents our analysis later from being biased by data "from the future." Since the data set is very clean and has few variables, no further data wrangling was required.

```{r}
#| label: training/test sets

sp_train <- SouthPark_IMDB_Data[which(SouthPark_IMDB_Data$Year <= 2013),]
sp_test <- SouthPark_IMDB_Data[which(SouthPark_IMDB_Data$Year > 2013),]
```

## Exploratory Data Analysis

In this section, my goal was to determine which variables in the data set would be most appropriate to include in my model.

First, I will explore the `AverageRating` variable, since this will be my response variable. This variable measures the average rating of an episode on IMDb on a scale from 1 to 10 stars. I assumed most of the ratings would come from people who had a strong opinion about the show. Furthermore, I believed that those who didn't like the show would stop watching after a few minutes, and thus not leave a rating. Therefore, I assumed that `AverageRating` would be centered around 8 stars with a relatively low standard deviation. I also thought that some episodes were bound to do poorly, so I assumed the distribution of `AverageRating` would have a left tail.

Looking at the 5 number summary as well as the mean and standard deviation of `AverageRating`, we can see that my assumptions about the center and variability were mostly correct. However, comparing the median to the mean, we can see that they are very similar, which suggests that the shape of the distribution is more symmetric than I thought. In fact, even the minimum is higher than I expected.

```{r}
#| label: AverageRating 5 number summary + mean & sd

sp_train |>
  summarize(
    min = min(AverageRating),
    Q1 = quantile(AverageRating, 0.25),
    median = median(AverageRating),
    Q3 = quantile(AverageRating, 0.75),
    max = max(AverageRating),
    mean = mean(AverageRating),
    sd = sd(AverageRating)
  )
```

Looking at the histogram for `AverageRating`, we can get a better idea of the shape of the distribution. While there is a slight left skew, the distribution is mostly symmetric.

```{r}
#| label: AverageRating histogram
#| message: FALSE

ggplot(
  data = sp_train,
  mapping = aes(
    x = AverageRating
  )
) +
  geom_histogram() +
  labs(
    x = "Average Rating out of 10",
    y = "Number of Episodes",
    title = "South Park IMDb"
  )
```

This begs the question, how does a show as controversial as South Park get such high average ratings? If we look at the IMDb page for South Park, we do see that it tends to get an overwhelming amount of high ratings compared to low ratings. Looking at other shows on IMDb, we see that they also tend to get higher average ratings in general, so this isn't something specific to South Park. It could be that users don't want to have disliked shows on their profile, or that they don't watch disliked episodes for long enough to give them a rating, but this is ultimately all speculation. However, it could be important to consider when using the average rating to measure the general reception of an episode that the information on IMDb comes from a voluntary sample. The average rating on IMDb might be higher than if we took a truly random sample of people who watched South Park and asked for their ratings.

Next, I moved on to the rest of the variables in the data set. I ignored the `PrimaryTitle` and `OriginalTitle` variables, since they basically serve as ID variables.

I suspected that the `Season` and `Year` variables might be highly correlated, since seasons are numbered in order of when they came out. To see if this was the case, I looked at a scatterplot for `Season` and `Year`. Each point represents one episode, although multiple points are stacked on top of each other because multiple episodes come out in the same season and year.

```{r}
#| label: Season & Year scatterplot

ggplot(
  data = sp_train,
  mapping = aes(
    x = Season,
    y = Year
  )
) +
  geom_point() +
  labs(
    title = "South Park IMDb"
  )
```

As we can see, `Season` and `Year` are almost perfectly linearly correlated, with only three of the earliest seasons spanning across two years. Thus, we don't want to use both of these variables in our modeling due to collinearity issues, which would lead to inflated standard errors and unreliable coefficient estimates. If these variables were to be included in the model, we would have to choose between one or the other, depending on which one seems more informative for our response. However, since the `Season` and `Year` variables serve to catalog when an episode was made, and I split my data based on `Year`, neither of these seemed appropriate to include in my model.

The `Episode` variable also serves to catalog when an episode was made, and I did not believe this would yield anything interesting when looked at on its own.

The histogram of `Runtime` doesn't yield anything surprising.

```{r}
#| label: Runtime histogram
#| message: FALSE

ggplot(
  data = sp_train,
  mapping = aes(
    x = Runtime
  )
) +
  geom_histogram() +
  labs(
    x = "Runtime (minutes)",
    y = "Number of Episodes",
    title = "South Park IMDb"
  )
```

An overwhelming majority of episodes have a runtime of about 22 minutes, with a few episodes being especially long. This is to be expected, since most television series have episodes that run for about 20 minutes. I thought the longer episodes might be TV specials, which run for longer than an average episode, but this does not seem to be the case. In fact, other than the unaired pilot having a 28 minute runtime, there does not seem to be anything special about the longer episodes, so it could simply be that these episodes just turned out longer. This means that for the majority of episodes, `Runtime` isn't a very informative variable. However, particularly long runtimes may have an effect on the average rating.

I thought that the `NumberOfVotes` variable would be the most informative variable, so I decided to explore it in more detail. Since South Park is a popular show, I expected `NumberOfVotes` to have a high center, somewhere in the thousands. Some episodes will be more popular than others, so I assumed there would be a high amount of variability.

Looking at the 5 number summary as well as the mean and standard deviation of `NumberOfVotes`, we can see that my guesses were correct. However, the mean is noticeably higher than the median, suggesting a right-skewed distribution.

```{r}
#| label: NumberOfVotes 5 number summary + mean & sd

sp_train |>
  summarize(
    min = min(NumberOfVotes),
    Q1 = quantile(NumberOfVotes, 0.25),
    median = median(NumberOfVotes),
    Q3 = quantile(NumberOfVotes, 0.75),
    max = max(NumberOfVotes),
    mean = mean(NumberOfVotes),
    sd = sd(NumberOfVotes)
  )
```

We have a much better idea of the extremeness of this skew when looking at the histogram.

```{r}
#| label: NumberOfVotes histogram
#| message: FALSE

ggplot(
  data = sp_train,
  mapping = aes(
    x = NumberOfVotes
  )
) +
  geom_histogram() +
  labs(
    x = "Number of Votes on IMDb",
    y = "Number of Episodes",
    title = "South Park IMDb"
  )
```

This skew seems to have been made even more extreme by a few outliers. This leads me to question, what are these points, and why do they have so many more votes?

The point on the far right corresponds to the episode "Scott Tenorman Must Die," which had one of the best receptions out of any South Park episode when it released. The other two rightmost points correspond to the episodes "Make Love, Not Warcraft" and "The Return of the Fellowship of the Ring to the Two Towers," whose popularity could be attributed to the obvious references in their titles attracting more viewers.

Another interesting point is the one on the far left, which corresponds to the unaired pilot. It makes sense that this wouldn't get as many votes, since the pilot of a show tends to be seen by less people. This is especially true for South Park's pilot, since it was never aired on television.

Looking at `AverageRating` and `NumberOfVotes` on their own made me wonder if there was a more direct connection between the average rating of a show and how many votes it got. I assumed that an episode having more votes meant it was more popular, and thus had a higher average rating.

We can look at a scatterplot between `NumberOfVotes` and `AverageRating`, where each point represents one episode.

```{r}
#| label: NumberOfVotes & AverageRating scatterplot

ggplot(
  data = sp_train,
  mapping = aes(
    x = NumberOfVotes,
    y = AverageRating
  )
) +
  geom_point() +
  labs(
    x = "Number of Votes on IMDb",
    y = "Average Rating out of 10",
    title = "South Park IMDb"
  )
```

My assumption generally appears to be true. 

Due to what was found earlier, I believe the relationship between `NumberOfVotes` and `AverageRating` would give me the most useful information compared to other relationships. Thus, I decided to follow up on the scatterplot between `NumberOfVotes` and `AverageRating`. Since the relationship appears nonlinear, we can plot a local regression curve to get a better idea of the relationship.

```{r}
#| label: NumberOfVotes & AverageRating scatterplot with loess
#| message: FALSE

ggplot(
  data = sp_train,
  mapping = aes(
    x = NumberOfVotes,
    y = AverageRating
  )
) +
  geom_point() +
  labs(
    x = "Number of Votes on IMDb",
    y = "Average Rating out of 10",
    title = "South Park IMDb"
  ) +
  geom_smooth(
    method = "loess",
    se = FALSE
  )
```

We can see that the point on the far left is dragging the left side of our curve up significantly. The question is, what should be done about this point?

Since this point has the least number of votes, we can deduce that it belongs to the unaired pilot. Since all future episodes of South Park will be aired, this point is probably not relevant to our analysis of future episodes. We might consider removing that point and only looking at episodes that were aired on television.

How does removing this point change our model? Below is a scatterplot with the point removed.

```{r}
#| label: NumberOfVotes & AverageRating scatterplot with pilot removed
#| message: FALSE

nopilot_train <- sp_train[-which(sp_train$PrimaryTitle == "Unaired Pilot"),]

ggplot(
  data = nopilot_train,
  mapping = aes(
    x = NumberOfVotes,
    y = AverageRating
  )
) +
  geom_point() +
  labs(
    x = "Number of Votes on IMDb",
    y = "Average Rating out of 10",
    title = "South Park IMDb without the Pilot"
  ) +
  geom_smooth(
    method = "loess",
    se = FALSE
  )
```

The local regression curve seems to be doing a better job of fitting the data on the left side of the graph now that the pilot was removed. The outliers on the right side of the graph pose no issue, since they follow the trend of the rest of the data.

For the remaining data, we might think that some episodes have less votes because they came later, and thus have been seen by less people. We can look at a scatterplot for `Season` and `NumberOfVotes` to see if this is the case.

```{r}
#| label: Season & NumberOfVotes scatterplot
#| message: FALSE

ggplot(
  data = nopilot_train,
  mapping = aes(
    x = Season,
    y = NumberOfVotes
  )
) +
  geom_point() +
  labs(
    x = "Season",
    y = "Number of Votes on IMDb",
    title = "South Park IMDb without the Pilot"
  )
```

While there may be a slightly downward trend, this does not seem to be the case in general.

Due to what I found, I decided to make my model using the training set with the pilot removed.

## Modeling

Here, I fit a k-nearest neighbors model predicting `AverageRating` from `Episode`, `Runtime`, and `NumberOfVotes`.

The way this model makes predictions for an observation is, for some positive integer k that we choose, it finds the k closest points in the training set to that observation. If our response is numerical, it predicts the average of those points' response values, and if our response is categorical, it predicts the most common class among those points. These points may also be weighted so that points that are closer to the observation have more influence over the predicted value or class of that observation. K-nearest neighbors does not make any assumptions.

When fitting the k-nearest neighbors model below, we will tune `neighbors`. This represents the "k" mentioned above, or the number of closest points to an observation that the model will consider when making predictions. For lower values of `neighbors`, the model will have higher variance and lower bias because the average response value (or most common class) will only be calculated using a few points (possibly even only one point), so predictions will vary greatly if a new point is added. For higher values of `neighbors`, the model will have higher bias and lower variance because the average response value (or most common class) will be calculated using the majority (if not all) of the training data, so adding a new point will not greatly influence the predictions.

```{r}
#| label: k-nearest neighbors model tuned
#| message: FALSE

knn_model <- nearest_neighbor(
  mode = "regression",
  neighbors = tune(),
  dist_power = 2
)

knn_recipe <- recipe(
  AverageRating ~ Episode + Runtime + NumberOfVotes,
  data = nopilot_train
) |>
  step_normalize(all_numeric_predictors())

knn_wflow <- workflow() |>
  add_model(knn_model) |>
  add_recipe(knn_recipe)

# cross validation
set.seed(437)
sp_kfold <- vfold_cv(nopilot_train, v = 5, repeats = 2)

knn_grid <- expand.grid(neighbors = seq(10, 100, by = 10))

knn_tune <- knn_wflow |>
  tune_grid(
    resamples = sp_kfold,
    grid = knn_grid,
    metrics = metric_set(rmse)
  )

knn_best <- select_best(knn_tune, metric = "rmse")

# fitting the model
knn_wflow_final <- knn_wflow |>
  finalize_workflow(parameters = knn_best)

knn_fit <- fit(knn_wflow_final, data = nopilot_train)

# avg rmse and standard error of avg rmse
knn_tune |>
  collect_metrics() |>
  filter(neighbors == knn_best$neighbors) |>
  dplyr::select(mean, std_err)
```

From the value of `neighbors` we chose in our cross validation, we see that we got an average RMSE of about 0.476 and a standard error of about 0.022. To know if this is good or not, let's compare it to what we would get with the null model (k-nearest neighbors model with 247 neighbors, the largest value of neighbors possible with our data).

```{r}
#| label: k-nearest neighbors model null
#| message: FALSE

# cross validation
knn_tune_null <- knn_wflow |>
  tune_grid(
    resamples = sp_kfold,
    grid = expand.grid(neighbors = 247),
    metrics = metric_set(rmse)
  )

# avg rmse and standard error of avg rmse
knn_tune_null |>
  collect_metrics() |>
  dplyr::select(mean, std_err)
```

We see that the null model gives us an average RMSE of about 0.552 and a standard error of about 0.020. The standard error is close to what we got with our tuned model, and the average RMSE is a bit higher. Thus, our model is performing slightly better than the null model.

## Conclusion

My goal was to predict the average rating of South Park episodes on IMDb. In my exploratory data analysis, I found that many of the variables in the data were not actually useful to include in my modeling. Variables that yielded interesting results were `NumberOfVotes`, which was found to have notable outliers, and `Runtime`, which was also found to have a few outliers with the majority of episodes having a runtime of 22 minutes. Using these variables as well as the `AverageRating` and `Episode` variables, I was able to fit a k-nearest neighbors model to predict the average rating of an episode. This model performed better than the null model.

### Limitations and Future Work

This k-nearest neighbors model falls flat in certain cases. We can look at the predictions on the test set sorted by highest to lowest residuals in absolute value to get a better understanding of this.

```{r}
#| label: episodes with most error
sp_preds <- augment(knn_fit, new_data = sp_test)
sp_preds |>
  mutate(
    abs_resid = abs(AverageRating - .pred)
  ) |>
  arrange(
    desc(abs_resid)
  ) |>
  select(PrimaryTitle, AverageRating, .pred, abs_resid)
```

As we can see, the model consistently makes poorer predictions on episodes with an average rating of less than 7 stars compared to other episodes. This makes sense considering that in the histogram of `AverageRating` seen in the exploratory data analysis, we saw that most episodes get an average rating above 7 stars.

```{r}
#| label: AverageRating histogram from EDA
#| message: FALSE
#| echo: FALSE

ggplot(
  data = sp_train,
  mapping = aes(
    x = AverageRating
  )
) +
  geom_histogram() +
  labs(
    x = "Average Rating out of 10",
    y = "Number of Episodes",
    title = "South Park IMDb"
  )
```

However, this means that if an episode has a low average rating, we would not be able to accurately predict that using this model. Thus, this model may not be useful in actually answering the question of whether or not an episode performed well among audiences.

Data that is collected through a truly random sample, rather than through a voluntary sample as is the case on IMDb, may be more useful in creating a model that predicts the performance of an episode. Furthermore, the fact that all of the data comes from the IMDb page for South Park means that in practice, a model such as the one shown here may not be useful. This is because if we need the information from IMDb to find the number of votes for an episode in order to estimate the average rating, we could also simply look up the average rating on IMDb.